---
title: "Time Series- Exploratory Data Analysis"
author: "Helen Yezerets"
date: "January 16, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploring raw time series
The most common first step when conducting time series analysis is to display your time series dataset in a visually intuitive format. The most useful way to view raw time series data in R is to use the print() command, which displays the Start, End, and Frequency of your data along with the observations.

Another useful command for viewing time series data in R is the length() function, which tells you the total number of observations in your data.

Some datasets are very long, and previewing a subset of data is more suitable than displaying the entire series. The head(___, n =___) and tail(___, n =___) functions, in which n is the number of items to display, focus on the first and last few elements of a given dataset respectively.

In this exercise, you'll explore the famous River Nile annual streamflow data, Nile. This time series dataset includes some metadata information. When calling print(Nile), note that Start = 1871 indicates that 1871 is the year of the first annual observation, and End = 1970 indicates 1970 is the year of the last annual observation.

```{r Nile, echo=FALSE}
print(Nile)
```

```{r echo=FALSE}
length(Nile)
# Display the first 10 elements of the Nile dataset

head(Nile, n=10)
# Display the last 12 elements of the Nile dataset
tail(Nile, n=12)
```
## Basic time series plots
While simple commands such as print(), length(), head(), and tail() provide crucial information about your time series data, another very useful way to explore any data is to generate a plot.

In this exercise, you will plot the River Nile annual streamflow data using the plot() function. For time series data objects such as Nile, a Time index for the horizontal axis is typically included. From the previous exercise, you know that this data spans from 1871 to 1970, and horizontal tick marks are labeled as such. The default label of "Time" is not very informative. Since these data are annual measurements, you should use the label "Year". While you're at it, you should change the vertical axis label to "River Volume (1e9 m^{3})".

Additionally, it helps to have an informative title, which can be set using the argument **main**. For your purposes, a useful title for this figure would be "Annual River Nile Volume at Aswan, 1871-1970".

Finally, the default plotting type for time series objects is **l** for line. Connecting consecutive observations can help make a time series plot more interpretable. Sometimes it is also useful to include both the observations points as well as the lines, and we instead use **b** for both.

```{r }
# Plot the Nile data
plot(Nile)


# Plot the Nile data with xlab, ylab, main, and type arguments
plot(Nile, xlab = "Year", ylab = "River Volume (1e9 m^{3})", 
main = "Annual River Nile Volume at Aswan, 1871-1970", type ="b")
```

##What does the time index tell us?
Some data are naturally evenly spaced by time. The time series **discrete_data** shown in the top figure has 20 observations, with one observation appearing at each of the discrete time indices 1 through 20. Discrete time indexing is appropriate for  **discrete_data**.

The time series **continuous_series** shown in the bottom figure also has 20 observations, it is following the same periodic pattern as **discrete_data**, but its observations are not evenly spaced. Its first, second, and last observations were observed at times 1.210322, 1.746137, and 20.180524, respectively. Continuous time indexing is natural for **continuous_series**, however, the observations are approximately evenly spaced, with about 1 observation observed per time unit. Let's investigate using a discrete time indexing for **continuous_series**.

```{r}
# Plot the continuous_series using continuous time indexing
par(mfrow=c(2,1))
#plot(continuous_time_index,continuous_series, type = "b")

# Make a discrete time index using 1:20 
discrete_time_index <-c(1:20)

# Now plot the continuous_series using discrete time indexing
#plot(discrete_time_index,continuous_series, type = "b")

```

##Identifying the sampling frequency
In addition to viewing your data and plotting over time, there are several additional operations that can be performed on time series datasets.

- start() and end() functions return the time index of the first and last observations, respectively. 

- time() function calculates a vector of time indices, with one element for each time index on which the series was observed.

- deltat() function returns the fixed time interval between observations 

- frequency() function returns the number of observations per unit time. 

- cycle() function returns the position in the cycle of each observation.

In this exercise, you'll practice applying these functions to the AirPassengers dataset, which reports the monthly total international airline passengers (in thousands) from 1949 to 1960.

```{r}
# Plot AirPassengers
plot(AirPassengers)
```

```{r}
# View the start and end dates of AirPassengers
start(AirPassengers)

end(AirPassengers)
# Use time(), deltat(), frequency(), and cycle() with AirPassengers 

```
#- time() function calculates a vector of time indices, with one element for each time index on which the series was observed.
```{r}
 time(AirPassengers)
```

#- deltat() function returns the fixed time interval between observations 

```{r}
deltat(AirPassengers)
```
###Frequency
```{r}
# - frequency() function returns the number of observations per unit time. 
 frequency(AirPassengers)
```

###Cycle
```{r}
# - cycle() function returns the position in the cycle of each observation.
 cycle(AirPassengers) 
 
```

##Missing values
Sometimes there are missing values in time series data, denoted NA in R, and it is useful to know their locations. It is also important to know how missing values are handled by various R functions. Sometimes we may want to ignore any missingness, but other times we may wish to **impute or estimate the missing values**.

Let's again consider the monthly AirPassengers dataset, but now the data for the year 1956 are missing. In this exercise, you'll explore the implications of this missing data and impute some new data to solve the problem.

- **mean()** function calculates the sample mean, but it fails in the presence of any NA values. Use **mean(___, na.rm = TRUE)** to calculate the mean with all missing values removed. It is common to replace missing values with the mean of the observed values. Does this simple data imputation scheme appear adequate when applied the the AirPassengers dataset?

```{r}
# Plot the AirPassengers data
plot(AirPassengers)

```

```{r}
# Compute the mean of AirPassengers
mean(AirPassengers, na.rm=TRUE)
```

```{r}
# Impute mean values to NA in AirPassengers
AirPassengers[85:96] <- mean(AirPassengers, na.rm = TRUE)
# Generate another plot of AirPassengers
plot(AirPassengers)
# Add the complete AirPassengers data to your plot
rm(AirPassengers)
points(AirPassengers, type = "l", col = 2, lty = 3)
```

##Creating a time series object with ts()
The function **ts()** can be applied to create time series objects. A time series object is a vector (univariate) or matrix (multivariate) with additional attributes, including time indices for each observation, the sampling frequency and time increment between observations, and the cycle length for periodic data. Such objects are of the ts class, and represent data that has been observed at (approximately) equally spaced time points. Now you will create time series objects yourself.

The advantage of creating and working with time series objects of the **ts** class is that many methods are available for utilizing time series attributes, such as time index information. For example, as you've seen in earlier exercises, calling **plot()** on a ts object will automatically generate a plot over time.

In this exercise, you'll familiarize yourself with the **ts** class by encoding some time series data (saved as data_vector) into ts and exploring the result. Your time series **data_vector** starts in the year 2004 and has 4 observations per year (i.e. it is quarterly data).

```{r}
# Use print() and plot() to view data_vector
data_vector<-c(10,6,11,8,10,3,6.9)

print(data_vector)

```
```{r}
plot(data_vector)
```
```{r}

# Convert data_vector to a ts object with start = 2004 and frequency = 4
time_series <- ts(data_vector, start=2004, frequency = 4)

# Use print() and plot() to view time_series
print(time_series)
```
```{r}
plot(time_series)

```

##Testing whether an object is a time series
When you work to create your own datasets, you can build them as ts objects. Recall the dataset data_vector from the previous exercise, which was just a vector of numbers, and time_series, the ts object you created from data_vector using the *ts()* function and information regarding the start time and the observation frequency. As a reminder, data_vector and time_series are shown in the plot on the right.

When you use datasets from others, such as those included in an R package, you can check whether they are ts objects using the *is.ts()* command. The result of the test is either TRUE when the data is of the ts class, or FALSE if it is not.

In this exercise, you'll explore the class of the datasets you've been using throughout this chapter.


```{r}
# Check whether data_vector and time_series are ts objects

is.ts(data_vector)
is.ts(time_series)

```

```{r}
# Check whether Nile is a ts object

is.ts(Nile)
# Check whether AirPassengers is a ts object
is.ts(AirPassengers)
```
##Plotting a time series object
It is often very useful to plot data we are analyzing, as is the case when conducting time series analysis. If the dataset under study is of the ts class, then the **plot()** function has methods that automatically incorporate time index information into a figure.

Let's consider the eu_stocks dataset (available in R by default as *EuStockMarkets*). This dataset contains daily closing prices of major European stock indices from 1991-1998, specifically, from Germany (DAX), Switzerland (SMI), France (CAC), and the UK (FTSE). The data were observed when the markets were open, so there are no observations on weekends and holidays. We will proceed with the approximation that this dataset has evenly spaced observations and is a four dimensional time series.

To conclude this chapter, this exercise asks you to apply several of the functions you've already learned to this new dataset.

```{r EuStockMarkets}


# Check whether eu_stocks is a ts object
is.ts(EuStockMarkets)

# View the start, end, and frequency of eu_stocks
eu_stocks<-EuStockMarkets
start(eu_stocks)
end(eu_stocks)
frequency(eu_stocks) 
```

```{r}

# Generate a simple plot of eu_stocks
plot(eu_stocks)

# Use ts.plot with eu_stocks
ts.plot(eu_stocks, col = 1:4, xlab = "Year", ylab = "Index Value", main = "Major European Stock Indices, 1991-1998")

```

## Introduction to time series data. Chapter 1 in Cowpertwait & Meltcalfe (2009)
###sections 1.1-1.4

###Recursive functions: Factorial 

```{r}
Fact <- function(n) if (n == 1) 1 else n * Fact(n - 1)
Fact(5)
```
```{r AirPassengers}
#data(AirPassengers)
AP <- AirPassengers
AP
```

```{r}
class(AP)
start(AP); end(AP); frequency(AP)
```
```{r}
summary(AP)
```

```{r}
plot(AP, ylab = "Passengers (1000's)")
```

## Seasonal Effect
To get a clearer view of the trend, the seasonal effect can be removed by
**aggregating the data to the annual level**, which can be achieved in R using the
*aggregate()* function. 
A **summary** of the values for each season can be viewed
using a boxplot, with the **cycle** function being used to *extract the seasons for each item of data*.
```{r}
layout(1:1)
plot(aggregate(AP))
boxplot(AP ~ cycle(AP))
```

```{r}
Maine.month <- read.delim("../Data/Maine.dat", header=TRUE)

#Maine.month <- read.table(www, header = TRUE)
attach(Maine.month)
class(Maine.month)
```

```{r}
Maine.month.ts <- ts(unemploy, start = c(1996, 1), frequency = 12)
```
This uses all the data. You can select a smaller number by specifying an
earlier end date using the parameter end. If we wish to analyse trends in the
unemployment rate, annual data will suffice. The average (mean) over the
twelve months of each year is another example of aggregated data, but this
time we divide by 12 to give a mean annual rate.

```{r}
Maine.annual.ts <- aggregate(Maine.month.ts)/12
```
We now plot both time series. There is clear monthly variation. From
Figure 1.3(a) it seems that the February figure is typically about 20% more
than the annual average, whereas the August figure tends to be roughly 20%
less.

```{r}
layout(1:2)
plot(Maine.month.ts, ylab = "unemployed (%)")
plot(Maine.annual.ts, ylab = "unemployed (%)")
```

We can calculate the precise percentages in R, using **window**. This function
will **extract that part of the time series between specified start and end points**
1.4 Plots, trends, and seasonal variation 9
and will sample with an interval equal to frequency if its argument is set to
TRUE. So, the first line below gives a time series of February figures.
```{r}

Maine.Feb <- window(Maine.month.ts, start = c(1996,2), freq = TRUE)
Maine.Aug <- window(Maine.month.ts, start = c(1996,8), freq = TRUE)
Feb.ratio <- mean(Maine.Feb) / mean(Maine.month.ts)
Aug.ratio <- mean(Maine.Aug) / mean(Maine.month.ts)


Maine.Feb
Maine.Aug
Feb.ratio
Aug.ratio
```
```{r}
layout(1:1)
plot(Maine.Feb)
plot(Maine.Aug)
```
```{r}
x <- c(1,2,3)

y <- matrix(c(3,1,2,2,1,3,3,2,1), ncol = 3, byrow = TRUE)
```

```{r}
y

```

```{r}
t(y)
```
```{r}

x%*%t(y)
```
```{r}
t <- seq(from = 1, to = 100, by = 1) + 10 + rnorm(100, sd = 7)
time_series<- ts(t, start=2004, frequency = 4)
plot(time_series)
```
```{r}
#www <- "http://www.massey.ac.nz/~pscowper/ts/USunemp.dat"
US.month <- read.delim("../Data/USunemp.dat", header=TRUE)

attach(US.month)
US.month.ts <- ts(US.month, start=c(1996,1), end=c(2006,10), frequency = 12)
plot(US.month.ts, ylab = "unemployed (%)")

```

The monthly supply of electricity (millions of kWh), beer (Ml),
and chocolate-based production (tonnes) in Australia over the period January
1958 to December 1990 are available from the Australian Bureau of Statistics
(ABS).

```{r}
#www <- "http://www.massey.ac.nz/~pscowper/ts/cbe.dat"
CBE <- read.delim("../Data/cbe.dat", header = T)
CBE[1:4, ]
```
```{r}
class(CBE)
```
create time series objects for the electricity, beer, and chocolate data.
If you omit end, R uses the full length of the vector, and if you omit the month
in start, R assumes 1. You can use plot with cbind to plot several series on
one figure (Fig. 1.5).

```{r}
Elec.ts <- ts(CBE[, 3], start = 1958, frequency =  12)
Beer.ts <- ts(CBE[, 2], start = 1958, frequency =  12)
Choc.ts <- ts(CBE[, 1], start = 1958, frequency =  12)
plot(cbind(Elec.ts, Beer.ts, Choc.ts), main="Chocolate, Beer, and Electricity Production: 1958???1990")
```
The intersection between the air passenger data and the electricity
data is obtained as follows


```{r}
AP.elec <- ts.intersect(AP, Elec.ts)
start(AP.elec)
#[1] 1958 1
```

```{r}
end(AP.elec)
#[1] 1960 12
```

```{r}
AP.elec[1:3, ]

```

```{r}
AP <- AP.elec[,1]; Elec <- AP.elec[,2]
layout(1:2)
plot(AP, main = "", ylab = "Air passengers / 1000's")
plot(Elec, main = "", ylab = "Electricity production / MkWh")
plot(as.vector(AP), as.vector(Elec),xlab = "Air passengers / 1000's",ylab = "Electricity production / MWh")
abline(reg = lm(Elec ~ AP))
```
##High correlation

```{r}
cor(AP, Elec)

```
In the plot function above, as.vector is needed to convert the ts objects to
ordinary vectors suitable for a scatter plot.

```{r}
#www <- "http://www.massey.ac.nz/~pscowper/ts/pounds_nz.dat"
Z <- read.delim("../Data/pounds_nz.dat", header = T)
Z[1:4, ]
#[1] 2.92 2.94 3.17 3.25

```

```{r}
Z.ts <- ts(Z, start = 1991, frequency = 4)
plot(Z.ts, xlab = "time / years",
ylab = "Quarterly exchange rate in $NZ / pound")
```
Two local trends are emphasised when the series is partitioned into two
subseries based on the periods 1992-1996 and 1996-1998. The window function
can be used to extract the subseries:


```{r}
Z.92.96 <- window(Z.ts, start = c(1992, 1), end = c(1996, 1))
Z.96.98 <- window(Z.ts, start = c(1996, 1), end = c(1998, 1))
layout (1:2)
plot(Z.92.96, ylab = "Exchange rate in $NZ/pound", xlab = "Time (years)" )
plot(Z.96.98, ylab = "Exchange rate in $NZ/pound", xlab = "Time (years)" )

```

In climate change studies (e.g., see Jones and Moberg, 2003; Rayner et al.
2003), the following global temperature series, expressed as anomalies from
the monthly means over the period 1961-1990, plays a central role:9

```{r}
#www <- "global.dat"
#Global <- scan(www)
#Global.ts <- ts(Global, start = c(1856, 1), end = c(2005, 12),frequency =  12)
Choc.annual <- aggregate(Choc.ts, FUN = mean)
plot(Choc.ts)
plot(Choc.annual)

```
using the
*time* function and the associated observed temperature series extracted using
window. The data are plotted and a line superimposed using a regression of
temperature on the new time index (Fig. 1.12).

```{r}
New.series <- window(Choc.ts, start=c(1970, 1), end=c(1985, 12))
New.time <- time(New.series)
plot(New.series); abline(reg=lm(New.series ~ New.time))
```