---
title: "Linear Regression"
author: "Helen Yezerets"
date: "February 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("devtools")
#devtools::install_github("nickpoison/astsa")
library(devtools)
library(astsa)
library(xts)

```
```{r}
data("jj")
head(jj)
```

```{r}

class(jj)
jj_xts<-as.xts(jj)
class(jj_xts)
```
```{r}
jj_xts
last(jj_xts, '1 year')
```

```{r}
#coredata(jj_xts)
#index(jj_xts)
periodicity(jj_xts)
start(jj)
end(jj)
frequency(jj)
plot(aggregate(jj), ylab = "Aggregated annually")
boxplot(jj ~ cycle(jj))
```

```{r}
data("sunspotz")
#head(sunspotz)
#class(sunspotz)
plot(sunspotz)
sun.decom<-decompose(sunspotz)
plot(sun.decom)
acf(sun.decom$random[2:458],lag.max=2, plot = FALSE)
```



```{r}
data("chicken")
summary(fit <- lm(chicken~time(chicken), na.action=NULL))
plot(chicken, ylab="cents per pound")
abline(fit) # add the fitted line
```

```{r}
culer = c(rgb(.9,.5,.5), rgb(.5,.9,.5), rgb(.5,.5,.9))
par(mfrow=c(3,1))
tsplot(cmort, main="Cardiovascular Mortality", col=culer[1], type="o", pch=19, ylab="")
tsplot(tempr, main="Temperature", col=culer[2], type="o", pch=19, ylab="")
tsplot(part, main="Particulates", col=culer[3], type="o", pch=19, ylab="")
```

```{r}
tsplot(cmort, main="", ylab="", ylim=c(20,130), col=culer[1])
lines(tempr, col=culer[2])
lines(part, col=culer[3])   
legend("topright", legend=c("Mortality", "Temperature", "Pollution"), lty=1, lwd=2, col=culer, bg="white")
##

```
```{r}
dev.new() # open a new graphic device
ts.plot(cmort,tempr,part, col=1:3) # all on same plot (not shown)
```


##Load dynlm
```{r}
fish = ts.intersect(rec, soiL6=lag(soi,-6), dframe=TRUE)
summary(fit1 <- lm(rec~soiL6, data=fish, na.action=NULL))
```

#The headache of aligning the lagged series can be avoided by using the R package `dynlm`, which must be downloaded and installed.
```{r}
library(dynlm)
summary(fit2 <- dynlm(rec~ L(soi,6)))

```
##Load Unemp data
##2. 
Reproduce the same workflow for unemp data (from astsa package)
```{r}
data("unemp")
summary(fit <- lm(unemp~time(unemp), na.action=NULL))
plot(unemp)
abline(fit) # add the fitted line
```

##3. R-squared and df, MSE,AIC and BIC
Follow steps from the example 2.2. Review stats in the chapter as needed.
* from summary(fit) obtain the results for R-squared and df
* from summary(aov(fit)) obtain MSE
* from AIC and BIC formula obtain AIC and BIC

Compare your results with the table 2.2 page 53.
```{r}
summary(fit)
summary(aov(fit))

```
##4. Reduce the model 
by removing one independent variable, starting from the last one:

cmort~ trend + temp + temp2 + part  becomes cmort~ trend + temp + temp2

Obtain the summary stats and compare with the table 2.2

```{r}
dev.new()
pairs(cbind(Mortality=cmort, Temperature=tempr, Particulates=part))
temp = tempr-mean(tempr) # center temperature
temp2 = temp^2
trend = time(cmort) # time
fit21 = lm(cmort~ trend + temp + temp2 + part, na.action=NULL)

fit20 = lm(cmort~ trend + temp + temp2, na.action=NULL)

fit19 = lm(cmort~ trend + temp, na.action=NULL)


```

##Summary statistics
```{r}
summary(fit21) # regression results
#R2=Multiple R-squared:  0.5954,	Adjusted R-squared:  0.5922 
summary(fit20)
summary(fit19)
```

##Compare ANOVA with Residuals
df, SSE=Sum Sq, MSE=Mean Sq
###fit21
```{r}

summary(aov(fit21)) # ANOVA table (compare to next line)

summary(aov(lm(cmort~cbind(trend, temp, temp2, part)))) # Table 2.1
```
###fit20
```{r}
#Residuals: df, SSE=Sum Sq, MSE=Mean Sq
summary(aov(fit20)) # ANOVA table (compare to next line)

#Residuals: df, SSE=Sum Sq, MSE=Mean Sq
summary(aov(lm(cmort~cbind(trend, temp, temp2)))) # Table 2.1
```
###fit19
```{r}
#Residuals: df, SSE=Sum Sq, MSE=Mean Sq
summary(aov(fit19)) # ANOVA table (compare to next line)

#Residuals: df, SSE=Sum Sq, MSE=Mean Sq
summary(aov(lm(cmort~cbind(trend, temp)))) # Table 2.1
```

##Obtain AIC and BIC
###fit21
```{r}
num = length(cmort) # sample size

#21
AIC(fit21)/num - log(2*pi) # AIC
BIC(fit21)/num - log(2*pi) # BIC
(AICc = log(sum(resid(fit21)^2)/num) + (num+5)/(num-5-2)) # AICc
```
###fit20
```{r}
#20
AIC(fit20)/num - log(2*pi) # AIC
BIC(fit20)/num - log(2*pi) # BIC
(AICc = log(sum(resid(fit20)^2)/num) + (num+5)/(num-5-2)) # AICc

```

###fit19
```{r}
#19
AIC(fit19)/num - log(2*pi) # AIC
BIC(fit19)/num - log(2*pi) # BIC
(AICc = log(sum(resid(fit19)^2)/num) + (num+5)/(num-5-2)) # AICc

```
##5. 
###fit18
Your last model should be cmort~trend
```{r}
fit18 = lm(cmort~ trend, na.action=NULL)
summary(fit18)
```


##Residuals: df, SSE=Sum Sq, MSE=Mean Sq
###fit18
```{r}
summary(aov(fit18)) # ANOVA table (compare to next line)

#Residuals: df, SSE=Sum Sq, MSE=Mean Sq
summary(aov(lm(cmort~cbind(trend)))) # Table 2.1
```
##AIC and BIC
###fit18
```{r}
#18
AIC(fit18)/num - log(2*pi) # AIC
BIC(fit18)/num - log(2*pi) # BIC
(AICc = log(sum(resid(fit18)^2)/num) + (num+5)/(num-5-2)) # AICc

```

##6
Load econ5 data
```{r}
data(econ5, package="astsa")
```
##7. 
Plot unemployment, gnp, consumption. Hint: you can extract each column with $. The following columns exist:  unemp, gnp, consum, govinv, prinv.

```{r}

par(mfrow=c(3,1))
tsplot(econ5$unemp, main="Unemployment",  col=culer[1],type="o",pch=19, ylab="")
tsplot(econ5$gnp, main="GNP", col=culer[2], type="o", pch=19, ylab="")
tsplot(econ5$consum, main="Consumption", col=culer[3], type="o", pch=19, ylab="")
```
##8. 
Create **pairs** with all three columns. Hint: you can extract each column with $

```{r}
dev.new()
pairs(cbind(Unemployment=econ5$unemp, GNP=econ5$gnp, GovInv=econ5$govinv, Consump= econ5$consum, Private=econ5$prinv))

```

##9.Convert to ts
econ5 is a dataframe. Convert it into a time series using **ts** function. 
Set start=c(1948, 3), end=c(1988, 2), frequency=4 - review econ5 information to understand why we are using frequency 4 (by quarter)

```{r}
econ5.ts <- ts(econ5, start=c(1948,3), end=c(1988,2), frequency = 4)
plot(econ5.ts)
```

##10. 
Create four variables for **lm()**. In the example below replace usdata with the name of your time series.
```{r}
trend = time(econ5.ts) #time
gnp <- econ5.ts[,2]
consum <- econ5.ts[,3]
govinv<-econ5.ts[,4]
response <- econ5.ts[,1]
```

##11. 
Create a model **lm()** with y=response and x1=trend, x2=gnp, x3=consum (independent variables). See step 4 above.
```{r}
fit_econ1 <- lm(response~ trend + gnp + consum, na.action=NULL)
fit_econ2 <- lm(response~ trend + gnp, na.action=NULL)
fit_econ3 <- lm(response~ trend, na.action=NULL)
#Create a linear model with unemployment, government investment and gnp.
fit_econ4<-lm(response~ trend + govinv + gnp, na.action=NULL)

```
##12
Obtain a summary of your model

```{r}
summary(fit_econ1)
plot(response)
abline(fit_econ1) # add the fitted line
```

```{r}

summary(fit_econ2)
plot(response)
abline(fit_econ2) # add the fitted line
```

```{r}
summary(fit_econ3)
plot(response)
abline(fit_econ3) # add the fitted line
```

##Gov Investment

```{r}
summary(fit_econ4)
plot(response)
abline(fit_econ4) # add the fitted line
```

