---
title: "Ch2_Covariance"
author: "Helen Yezerets"
date: "February 9, 2019"
output: html_document

references: Cowpertwait, P. S. P., & Metcalfe, A. V. (2009). Introductory Time Series with R. (Ch.2) New York: Springer. https://doi.org/10.1007/978-0-387-88698-5. 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
data(AirPassengers)
AirPassengers
```
```{r}
AP <- AirPassengers
AP.decom <- decompose(AP, "multiplicative") #makes seasonal adjustments using moving average, depicts original ts, trend, seasonal, and random (remainder) part
plot(ts(AP.decom$random[7:138]))
acf(AP.decom$random[7:138]) #plots autocovariance or autocorrelation


```

###Cosine shape is a characteristic of autoregressive model of order 2 or  a failed seasonal adjustment by **decompose()** function. We need to check the st dev

```{r}
sd(AP[7:138])
sd(AP[7:138] - AP.decom$trend[7:138])
sd(AP.decom$random[7:138])
##The reduction in the standard deviation shows that the seasonal adjustment has been very effective.
```

##Read and plot the data
```{r}
Fontdsdt.dat<- read.table("../Data/Fontdsdt.dat", header = TRUE)
attach(Fontdsdt.dat)
plot(ts(adflow), ylab = 'adflow')
acf(adflow, xlab = 'lag (months)', main="")

```
##Compute correlation:
###Variant 1
```{r}
cov(x,y)/(sd(x)*sd(y))

```
###Variant 2
```{r}
cor(x,y)
```
- regression approach  assumes a linear trend
- decompose smooths the time series without assuming any particular form for the trend.

###2.5 Summary of commands used in examples
```{r ignore=TRUE}

mean(x) # returns the mean (average)
var(x) #returns the variance
sd(x) #returns the standard deviation
cov(x,y) #returns the covariance 
cor(x,y) #returns the correlation
acf(y, lag.max=10,plot=FALSE) #returns the correlogram (or sets the argument to obtain autocovariance function)

```
```{r}
Fontdsdt.dat<- read.table("../Data/Fontdsdt.dat", header = TRUE)
attach(Fontdsdt.dat)
plot(ts(adflow), ylab = 'adflow')
acf(adflow, xlab = 'lag (months)', main="")

```
Compute correlation:
Variant 1
```{r}
cov(x,y)/(sd(x)*sd(y))

```
Variant 2
```{r}
cor(x,y)
```

The main difference between the regression approach and using
decompose is that the former assumes a linear trend, whereas the latter
smooths the time series without assuming any particular form for the trend.

```{r}
Varnish<- read.table("../Data/varnish.dat", header = TRUE)
#1) the mean for y,  
mean(Varnish$y)
#2) the standard deviation for y,
sd(Varnish$y)
# 3) covariance between x and y.
cov(Varnish$x,Varnish$y)
```
#Import guessWhat data into R, Draw a scatterplot, calculate a correlation between x and y, plot the data

```{r}
guessWhat<-read.table("../Data/guesswhat.dat", header = TRUE)
cor(guessWhat$x, guessWhat$y)
cov(guessWhat$x, guessWhat$y)

plot(guessWhat)
```

##Import cbe data
```{r}
cbe <- read.table("../Data/cbe.dat", header = TRUE)

# Create a time series choc.ts for the first column [chocolate production]. Starting point is 1958, frequency - 12. 
Elec.ts <- ts(cbe[, 3], start = 1958, frequency = 12)
Choc.ts <- ts(cbe[, 1], start = 1958, frequency = 12)
```
# Plot choc.ts.  
```{r}
plot(cbind(Elec.ts,Choc.ts))
```
##Decide if your model should be additive or multiplicative. Hint: compare with electricity data, does the variance of the chocolate data increase with the trend increase?
```{r}
plot(decompose(Choc.ts))
Choc.decom <- decompose(Choc.ts, type="additive")
plot(Choc.decom)
Trend <- Choc.decom$trend
Seasonal <- Choc.decom$seasonal
ts.plot(cbind(Trend, Trend + Seasonal), lty = 1:2)
```


```{r}
Choc.ts1 <- ts(cbe[, 1], start = 1958, frequency = 12)

#Calculate standard deviation for chocolate time series (not decomposed)
sd(Choc.ts1) 
Choc.decomp1<- decompose(Choc.ts1, type="additive")

#Calculate standard deviation for chocolate time series [12 indices removed] subtracting trend [12 indices removed]
sd(Choc.decomp1$x[7:390]-Choc.decomp1$trend[7:390])

#Calculate standard deviation after seasonal adjustments [12 indices removed]
sd(Choc.decomp1$x[7:390]-Choc.decomp1$seasonal[7:390])

```

1771 1051 1425
```{r}
acf(Choc.decom$random[7:390])
```
