---
title: "datacamp_TS3_Scatterplots"
author: "Helen Yezerets"
date: "January 30, 2019"
output: html_document
tags: scatterplot, covariance, correlation, autocorrelation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Asset prices vs. asset returns

Log returns, also called continuously compounded returns, are also commonly used in financial time series analysis. They are the log of gross returns, or equivalently, the changes (or first differences) in the logarithm of prices.

One advantage of using log returns is that calculating multi-period returns from individual periods is greatly simplified - you just add them together!
```{r EuStockMarkets}
eu_stocks<-EuStockMarkets

# Plot eu_stocks
plot(eu_stocks)

# Use this code to convert prices to returns
returns <- eu_stocks[-1,] / eu_stocks[-1860,] - 1

# Convert returns to ts
returns <- ts(returns, start = c(1991, 130), frequency = 260)

# Plot returns
plot(returns)

# Use this code to convert prices to log returns
logreturns <- diff(log(eu_stocks))

# Plot logreturns
plot(logreturns)

```
Daily net returns and daily log returns are two valuable metrics for financial data.

##Characteristics of financial time series
Daily financial asset returns typically share many characteristics. Returns over one day are typically small, and their average is close to zero. At the same time, their variances and standard deviations can be relatively large. Over the course of a few years, several very large returns (in magnitude) are typically observed. These relative outliers happen on only a handful of days, but they account for the most substantial movements in asset prices. Because of these extreme returns, the distribution of daily asset returns is not normal, but heavy-tailed, and sometimes skewed. In general, individual stock returns typically have even greater variability and more extreme observations than index returns.

eu_percentreturns dataset is the percentage returns calculated from eu_stocks data. 
Notice that the average daily return is about 0, while the standard deviation is about 1 percentage point. Also apply the **hist() and qqnorm()** functions to make histograms and normal quantile plots, respectively, for each of the indices.

```{r eu_percentreturns}

eu_percentreturns<-logreturns
# Generate means from eu_percentreturns
colMeans(eu_percentreturns)

# Use apply to calculate sample variance from eu_percentreturns
apply(eu_percentreturns, MARGIN = 2, FUN = var)

# Use apply to calculate standard deviation from eu_percentreturns
apply(eu_percentreturns, MARGIN = 2, FUN = sd)

# Display a histogram of percent returns for each index
par(mfrow = c(2,2))
apply(eu_percentreturns, MARGIN = 2, FUN = hist, main = "Percent returns ", xlab = "Percentage Return")

# Display normal quantile plots of percent returns for each index
par(mfrow = c(2,2))
apply(eu_percentreturns, MARGIN = 2, FUN = qqnorm, main = "Normal Quantile for each index", xlab = "Percentage Return")

qqline(eu_percentreturns)

```

##Plotting pairs of data
eu_stocks contains daily closing prices from 1991-1998 for the major stock indices in Germany (DAX), Switzerland (SMI), France (CAC), and the UK (FTSE).

to examine the bivariate relationship between pairs of time series. The *plot(a, b)* function will produce a scatterplot when two time series names a and b are given as input.

To simultaneously make scatterplots for all pairs of several assets the `pairs()` function can be applied to produce a *scatterplot matrix*. When shared time trends are present in prices or index values it is common to instead compare their returns or log returns.

Note that the normal distribution has elliptical contours of equal probability, and pairs of data drawn from the multivariate normal distribution form a roughly elliptically shaped point cloud. 
```{r}
#eu_stocks
# Make a scatterplot of DAX and FTSE
plot(eu_stocks)

# Make a scatterplot matrix of eu_stocks
pairs(eu_stocks)

# Convert eu_stocks to log returns
logreturns <- diff(log(eu_stocks))

# Plot logreturns
plot(logreturns)

# Make a scatterplot matrix of logreturns
pairs(logreturns)

```
the `pairs()` command is a useful way to quickly check for relationships between your indices.

##Calculating sample covariances and correlations
**Sample covariances** measure the strength of the linear relationship between matched pairs of variables. The `cov()` function can be used to calculate covariances for a pair of variables, or a covariance matrix when a matrix containing several variables is given as input. For the latter case, the matrix is symmetric with covariances between variables on the off-diagonal and variances of the variables along the diagonal. 

Covariances are very important throughout finance, but they are *not scale free* and they can be *difficult to directly interpret*. 
**Correlation** is the standardized version of covariance that ranges in value from -1 to 1, where values close to 1 in magnitude indicate a strong linear relationship between pairs of variables. The `cor()` function can be applied to both pairs of variables as well as a matrix containing several variables, and the output is interpreted analogously.

```{r}
head(logreturns)
DAX_logreturns<-logreturns[, "DAX"]
FTSE_logreturns<-logreturns[, "FTSE"]

```
```{r}
# Use cov() with DAX_logreturns and FTSE_logreturns
cov(DAX_logreturns,FTSE_logreturns )

# Use cov() with logreturns
cov(logreturns)

# Use cor() with DAX_logreturns and FTSE_logreturns
cor(DAX_logreturns, FTSE_logreturns)

# Use cor() with logreturns
cor(logreturns)
```
##Calculating autocorrelations
Autocorrelations or lagged correlations are used to assess whether a time series is dependent on its past. For a time series x of length n we consider the n-1 pairs of observations one time unit apart. The first such pair is (x[2],x[1]), and the next is (x[3],x[2]). Each such pair is of the form (x[t],x[t-1]) where t is the observation index, which we vary from 2 to n in this case. The lag-1 autocorrelation of x can be estimated as the sample correlation of these (x[t], x[t-1]) pairs.

In general, we can manually create these pairs of observations. First, create two vectors, x_t0 and x_t1, each with length n-1, such that the rows correspond to (x[t], x[t-1]) pairs. Then apply the cor() function to estimate the lag-1 autocorrelation.

Luckily, the `acf()` command provides a shortcut. Applying `acf(..., lag.max = 1, plot = FALSE)` to a series x automatically calculates the lag-1 autocorrelation.

Finally, note that the two estimates differ slightly as they use slightly different scalings in their calculation of sample covariance, 1/(n-1) versus 1/n. Although the latter would provide a biased estimate, it is preferred in time series analysis, and the resulting autocorrelation estimates only differ by a factor of (n-1)/n.

```{r}
set.seed(123)
n=length(DAX_logreturns)
x<-ts(DAX_logreturns)

# Define x_t0 as x[-1]
x_t0 <- x[-1]

# Define x_t1 as x[-n]
x_t1 <- x[-n]

# Confirm that x_t0 and x_t1 are (x[t], x[t-1]) pairs  
head(cbind(x_t0, x_t1))
  
# Plot x_t0 and x_t1
plot(x_t0, x_t1)
```

```{r}

# View the correlation between x_t0 and x_t1
cor(x_t0, x_t1)

# Use acf with x
acf(x, lag.max = 1, plot = FALSE)

# Confirm that difference factor is (n-1)/n
cor(x_t1, x_t0) * (n-1)/n

```
##The autocorrelation function
Autocorrelations can be estimated at many lags to better assess how a time series relates to its past. We are typically most interested in how a series relates to its most recent past.

The `acf(..., lag.max = ..., plot = FALSE)` function will estimate all autocorrelations from 0, 1, 2,..., up to the value specified by the argument lag.max. 

```{r}
# Generate ACF estimates for x up to lag-10
acf(x, lag.max = 10, plot = FALSE)

# Type the ACF estimate at lag-10 
0.009

# Type the ACF estimate at lag-5
-0.032

```
 Since autocorrelations may vary by lag, we often consider **autocorrelations as a function of the time lag**. Taking this view, we have now estimated the autocorrelation function (ACF) of x from lags 0 to 10.
 
##Visualizing the autocorrelation function
Estimating the autocorrelation function (ACF) at many lags allows us to assess how a time series x relates to its past. The numeric estimates are important for detailed calculations, but it is also useful to visualize the ACF as a function of the lag.

In fact, the `acf()` command produces a figure by default. It also makes a default choice for lag.max, the maximum number of lags to be displayed.

 **strong persistence**, meaning the current value is closely related to those that proceed it. 
**periodic pattern with a cycle length of approximately four observations**, meaning the current value is relatively close to the observation four before it. 
does not exhibit any clear pattern.

In the plots produced by acf(), the **lag** for each autocorrelation estimate is denoted on the **horizontal axis** and each **autocorrelation estimate** is indicated by the height of the **vertical** bars. Recall that the ACF at lag-0 is always 1.

Finally, each ACF figure includes a pair of blue, horizontal, dashed lines representing lag-wise 95% confidence intervals centered at zero. These are used for determining the statistical significance of an individual autocorrelation estimate at a given lag versus a null value of zero, i.e., no autocorrelation at that lag.

```{r}
# View the ACF of x
acf(x)

```

```{r}

y<-ts(sample(FTSE_logreturns, size = n))
# View the ACF of y
acf(y)

```

```{r}

z<-ts(sample(logreturns[, "SMI"], size = n))
# View the ACF of z
acf(z)

```
The estimated ACF is near zero at all lags. It appears the series x, y, or z are **not linearly related to its past**, at least through lag 35.

##References
Matteson, D. S. (2019). *Introduction to Time Series Analysis*. Retrieved from https://www.datacamp.com/courses/introduction-to-time-series-analysis
