---
title: "TS_StohasticProcess"
author: "Helen Yezerets"
date: "February 2, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
install.packages("astsa")
library(astsa) 
plot(jj, type="o", ylab="Quarterly Earnings per Share")

```


```{r}
library(TTR)
djia = getYahooData("^DJI", start=20060420, end=20160420, freq="daily")
library(xts)
djiar = diff(log(djia$Close))[-1] # approximate returns
plot(djiar, main="DJIA Returns", type="n")
lines(djiar)
```

```{r}
crazy <- function(num) {
  x <- c() #placeholder for a vector
  for (n in 1:num) { x[n] <- mean(rcauchy(n)) } # creates vector of means of random Cauchy variates 
  plot(x, type="l", xlab="sample size", ylab="sample mean")
}

crazy(200) #call function crazy

```
##R.4 Time Series Primer
```{r}
#install.packages("astsa")
#library(astsa)
```

To create a time series object, use the command **ts**. Related commands are **as.ts** to coerce an object to a time series and **is.ts** to test whether an object is a time series. First, make a small data set:
```{r}
(mydata = c(1,2,3,2,1) ) # make it and view it
```
```{r}
#Now make it a time series:
(mydata = as.ts(mydata) )
```

Make it an annual time series that starts in 1950:
```{r}
(mydata = ts(mydata, start=1950) )
```

Now make it a quarterly time series that starts in 1950-III:
```{r}
(mydata = ts(mydata, start=c(1950,3), frequency=4) )
```
```{r}
time(mydata) # view the sampled times
```

To use part of a time series object, use window():
```{r}
(x = window(mydata, start=c(1951,1), end=c(1951,3) ))
```

Next, we'll look at lagging and differencing. 
First make a simple series, xt :
```{r}
x = ts(1:5)
```
Now, column bind (cbind) lagged values of xt and you will notice that lag(x) is
forward lag, whereas lag(x, -1) is backward lag.

```{r}
cbind(x, lag(x), lag(x,-1))
```

Compare cbind and ts.intersect:
```{r}
ts.intersect(x, lag(x,1), lag(x,-1))
```

To difference a series, rxt = xt ???? xt????1, use
diff(x) but note that diff(x, 2)
is not second order differencing,  do one of these:

```{r}
diff(diff(x))
diff(x, diff=2) # same thing
```
and so on for higher order differencing.
We will also make use of regression via lm(). First, suppose we want to fit a
simple linear regression, y =  alfa+ beta*x + epsilon. In R, the formula is written as y~x:
```{r}
set.seed(1999)
x = rnorm(10)
y = x + rnorm(10)
summary(fit <- lm(y~x) )
```

```{r}
plot(x, y) # draw a scatterplot of the data 
abline(fit) # add the fitted line to the plot 
```

##All sorts of information can be extracted from the lm object, which we called fit.
For example,
```{r}
resid(fit) # will display the residuals 
fitted(fit) # will display the fitted values
lm(y ~ 0 + x) # will exclude the intercept 
```
You have to be careful if you use lm() for lagged values of a time series. If you use lm(), then what you have to do is align the series using ts.intersect.
Please read the warning Using time series in the lm() help file [help(lm)]. Here
is an example regressing astsa data, weekly cardiovascular mortality (cmort) on
we create ded, which consists of the intersection of the three series:
```{r}
ded = ts.intersect(cmort, part, part4=lag(part,-4))
```
Now the series are all aligned and the regression will work.
```{r}
summary(fit <- lm(cmort~part+part4, data=ded, na.action=NULL) )
```

There was no need to rename lag(part,-4) to part4, it's just an example of what
you can do.
An alternative to the above is the package **dynlm**, which has to be installed. After
the package is installed, the previous example may be run as follows:
```{r}
#install.packages("dynlm")
library(dynlm) # load the package
fit = dynlm(cmort~part + L(part,4)) # no new data file needed
summary(fit)
```
The output is identical to the lm output. To fit another model, for example, add the
temperature series tempr, the advantage of dynlm is that a new data file does not have
to be created. We could just run
```{r}
summary(dynlm(cmort~ tempr + part + L(part,4)) )
```
```{r}
#Regression model for JJ
trend = time(jj) - 1970 # helps to 'center' time
Q = factor(cycle(jj) ) # make (Q)uarter factors
reg = lm(log(jj)~0 + trend + Q, na.action=NULL) # no intercept
model.matrix(reg) # view the model design matrix

```
```{r}
summary(reg) # view the results (not shown)
```
The workhorse for ARIMA simulations is arima.sim. Here are some examples;
```{r}
x = arima.sim(list(order=c(1,0,0), ar=.9), n=100) + 50 # AR(1) w/mean 50
x = arima.sim(list(order=c(2,0,0), ar=c(1,-.9)), n=100) # AR(2)
x = arima.sim(list(order=c(1,1,1), ar=.9 ,ma=-.5), n=200) # ARIMA(1,1,1)
```
Some oscillations are taken out. A linear combination of values in a time series such as in (1.1) is referred to,generically, as a **filtered series**; hence the command filter in the following code
```{r}
w = rnorm(500,0,1) # 500 N(0,1) variates
v = filter(w, sides=2, filter=rep(1/3,3)) # moving average
par(mfrow=c(2,1))
plot.ts(w, main="white noise")
plot.ts(v, ylim=c(-3,3), main="moving average")
```

```{r}
set.seed(154) # so you can reproduce the results
w = rnorm(200); x = cumsum(w) # two commands in one line
wd = w +.2; xd = cumsum(wd)
plot.ts(xd, ylim=c(-5,55), main="random walk", ylab='')
lines(x, col=4); abline(h=0, col=4, lty=2); abline(a=0, b=.2, lty=2)
```
```{r faithful}
data(faithful)
(FT<-faithful)
x<-FT$eruptions
y<-FT$waiting
cov(x,y)
```

```{r}
cov(x, y,  method = c("pearson", "kendall", "spearman"))
```

```{r}
cor(x, y,    method = c("pearson", "kendall", "spearman"))

```

#Referencecs: Shumway, R., & Stooffer, D. S. (2016). Time Series Analysis and Its Applications With R Examples (Fourth Edition). Springer.