---
title: "Week2 Time Series Cowpertwait Chapter 1"
author: "Olga Scrivner"
date: "1/20/2019"
output: html_document

references: Cowpertwait, P. S. P., & Metcalfe, A. V. (2009). Introductory Time Series with R. (Ch.1) New York: Springer. https://doi.org/10.1007/978-0-387-88698-5. 
---

## Introduction: Plots, trends
AirPassenger data set is a sample time series data - the number of air passengers booking for the period 1949-1960.
```{r AirPassenger}
AP <- AirPassengers
AP
```

### Explore the data set
```{r}
class(AP)
start(AP)
end(AP)
frequency(AP)
```
##Preliminary plot
```{r message=FALSE, echo=FALSE}
plot(AP, ylab = "Passengers (1000's)")
```

##Reduce seasonal effect by aggregating data to the annual level using aggregate() and cycle() . 
```{r message=FALSE, echo=FALSE}
layout(1:2)
plot(aggregate(AP), ylab = "Aggregated annually")
boxplot(AP ~ cycle(AP))
```
From boxplot - most people travel from June-September

## Unemployment Data
```{r}
Maine.month <- read.table("../Data/Maine.dat", header=TRUE)
attach(Maine.month)
class(Maine.month)
head(Maine.month)
```

Use `ts` to convert to time series
```{r}
Maine.month.ts <- ts(unemploy, start = c(1996, 1), frequency = 12)
```

##Calculate the annual mean

```{r  message=FALSE, echo=FALSE}
Maine.annual.ts <- aggregate(Maine.month.ts)/12

plot(Maine.month.ts, ylab = "unemployed (%)", main="Monthly:January 1996-August 2006")

plot(Maine.annual.ts, ylab = "unemployed (%)", main="Annual: 1996-2006")

```
##Calculate precise percentage for unemployment

```{r}
Maine.Feb <- window(Maine.month.ts, start = c(1996,2), freq = TRUE)
Maine.Aug <- window(Maine.month.ts, start = c(1996,8), freq = TRUE)
Feb.ratio <- mean(Maine.Feb) / mean(Maine.month.ts)
Aug.ratio <- mean(Maine.Aug) / mean(Maine.month.ts)
```
February ration is `r Feb.ratio` and August ration is `r Aug.ratio`

## Multiple Time Series
```{r echo=FALSE}
cbe <- read.table("../Data/cbe.dat", header = TRUE)
head(cbe)
```
##Create a time series for each column:
```{r echo=FALSE}
Elec.ts <- ts(cbe[, 3], start = 1958, frequency = 12)
Beer.ts <- ts(cbe[, 2], start = 1958, frequency = 12)
Choc.ts <- ts(cbe[, 1], start = 1958, frequency = 12)
```
##Plot all three using cbind():
```{r echo=FALSE}
plot(cbind(Elec.ts, Beer.ts, Choc.ts))
```

##Use intersect function and plot
```{r}
AP.elec <- ts.intersect(AP, Elec.ts)
AP <- AP.elec[,1]
Elec <- AP.elec[,2]
layout(1:3) 
  
plot(AP, main = "", ylab = "Air passengers / 1000's") 
plot(Elec, main = "", ylab = "Electricity production / MkWh")
plot(as.vector(AP), as.vector(Elec), xlab = "Air passengers / 1000's", ylab = "Electricity production / MWh")
abline(reg = lm(Elec ~ AP))
```

