---
title: "Time Series- Exploratory Data Analysis"
author: "Helen Yezerets"
date: "January 16, 2019"
output: html_document


references:
Matteson, D. S. (2019). Introduction to Time Series Analysis (Ch.1). Retrieved from https://www.datacamp.com/courses/introduction-to-time-series-analysis

Cowpertwait, P.S.P., & Metcalfe, A.V. (2009). Introductory Time Series with R. New York: Springer. https://doi.org/10.1007/978-0-387-88698-5 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Datacamp *Introduction to Time Series Analysis*
## Exploring raw time series
To view raw time series data in R we can use the `print()` command, which displays the Start, End, and Frequency of your data along with the observations.

`length()` function tells the total number of observations in your data.

Previewing a subset of data is more suitable than displaying the entire series. The `head(___, n =___)` and `tail(___, n =___)` functions, in which **n is the number of items to display**, focus on the first and last few elements of a given dataset respectively.

```{r Nile, echo=FALSE}
print(Nile)
length(Nile)
# Display the first 10 elements of the Nile dataset
head(Nile, n=10)
# Display the last 12 elements of the Nile dataset
tail(Nile, n=12)
```
## Basic time series plots
`plot()` function:
to have an informative title, which can be set using the argument **main**. 
Example:main= "Annual River Nile Volume at Aswan, 1871-1970".

Finally, the default plotting type for time series objects is **l** for line. Connecting consecutive observations can help make a time series plot more interpretable. Sometimes it is also useful to include both the observations points as well as the lines, and we instead use **b** for both.

```{r }
# Plot the Nile data
plot(Nile)

# Plot the Nile data with xlab, ylab, main, and type arguments
plot(Nile, xlab = "Year", ylab = "River Volume (1e9 m^{3})", 
main = "Annual River Nile Volume at Aswan, 1871-1970", type ="b")
```

##What does the time index tell us?
Some data are naturally evenly spaced by time. The time series **discrete_data** shown in the top figure has 20 observations, with one observation appearing at each of the discrete time indices 1 through 20. Discrete time indexing is appropriate for  **discrete_data**.

The time series **continuous_series** shown in the bottom figure also has 20 observations, it is following the same periodic pattern as **discrete_data**, but its observations are not evenly spaced. Its first, second, and last observations were observed at times 1.210322, 1.746137, and 20.180524, respectively. Continuous time indexing is natural for **continuous_series**, however, the observations are approximately evenly spaced, with about 1 observation observed per time unit. Let's investigate using a discrete time indexing for **continuous_series**.

```{r}
# Plot the continuous_series using continuous time indexing
par(mfrow=c(2,1))
#plot(continuous_time_index,continuous_series, type = "b")
# discrete time index using 1:20 
discrete_time_index <-c(1:20)

# Now plot the continuous_series using discrete time indexing
#plot(discrete_time_index,continuous_series, type = "b")

```

##Identifying the sampling frequency
- `start() and end()` functions return the time index of the first and last observations, respectively. 

- `time()`function calculates a **vector of time indices**, with one element for each time index on which the series was observed.

- `deltat()` function returns the **fixed time interval between observations** 

- `frequency()` function returns the **number of observations per unit time**. 

- `cycle()` function returns the **position in the cycle** of each observation.

```{r}
# Plot AirPassengers
plot(AirPassengers)
```

```{r}
# View the start and end dates of AirPassengers
start(AirPassengers)

end(AirPassengers)
```
### time() function calculates a vector of time indices, with one element for each time index on which the series was observed.
```{r}
 time(AirPassengers)
```

### deltat() function returns the fixed time interval between observations 

```{r}
deltat(AirPassengers)
```
### frequency() function returns the number of observations per unit time. 
```{r}

 frequency(AirPassengers)
```

### cycle() function returns the position in the cycle of each observation.
```{r}

 cycle(AirPassengers) 
 
```

##Missing values
Sometimes there are missing values in time series data, denoted NA in R, and it is useful to know their locations. It is also important to know how missing values are handled by various R functions. Sometimes we may want to ignore any missingness, but other times we may wish to **impute or estimate the missing values**.

`mean()` function calculates the sample mean, but it fails in the presence of any NA values. Use `mean(___, na.rm = TRUE)` to calculate the mean with all missing values removed. It is common to **replace missing values with the mean** of the observed values. 

```{r}
# Plot the AirPassengers data
plot(AirPassengers)

```

```{r}
# Compute the mean of AirPassengers
mean(AirPassengers, na.rm=TRUE)
```

```{r}
# Impute mean values to NA in AirPassengers
AirPassengers[85:96] <- mean(AirPassengers, na.rm = TRUE)
# Generate another plot of AirPassengers
plot(AirPassengers)
# Add the complete AirPassengers data to your plot
rm(AirPassengers)
points(AirPassengers, type = "l", col = 2, lty = 3)
```

##Creating a time series object with ts()
The function `ts()` is used to create time series objects. A time series object is a vector (univariate) or matrix (multivariate) with additional attributes, including time indices for each observation, the sampling frequency and time increment between observations, and the cycle length for periodic data. Such objects are of the ts class, and represent data that has been observed at **(approximately) equally spaced time points**. 

Calling `plot()` on a ts object will automatically generate a plot over time.
```{r}
# Use print() and plot() to view data_vector
data_vector<-c(10,6,11,8,10,3,6.9)

print(data_vector)

```
```{r}
plot(data_vector)
```
```{r}

# Convert data_vector to a ts object with start = 2004 and frequency = 4
time_series <- ts(data_vector, start=2004, frequency = 4)

# Use print() and plot() to view time_series
print(time_series)
```
```{r}
plot(time_series)

```

##Testing whether an object is a time series
To check whether the objects are ts objects use the `is.ts()` command. The result of the test is either TRUE when the data is of the ts class, or FALSE if it is not.

```{r}
# Check whether data_vector and time_series are ts objects

is.ts(data_vector)
is.ts(time_series)
```

```{r}
# Check whether Nile is a ts object
is.ts(Nile)
# Check whether AirPassengers is a ts object
is.ts(AirPassengers)
```
##Plotting a time series object
`plot()` function has methods that automatically incorporate time index information into a figure.

```{r EuStockMarkets}
# Check whether eu_stocks is a ts object
is.ts(EuStockMarkets)

# View the start, end, and frequency of eu_stocks
eu_stocks<-EuStockMarkets
start(eu_stocks)
end(eu_stocks)
frequency(eu_stocks) 
```

```{r}

# Generate a simple plot of eu_stocks
plot(eu_stocks)

# Use ts.plot with eu_stocks
ts.plot(eu_stocks, col = 1:4, xlab = "Year", ylab = "Index Value", main = "Major European Stock Indices, 1991-1998")

```

## Introduction to time series data. Chapter 1 in Cowpertwait & Meltcalfe (2009)
###sections 1.1-1.4
###Summary functions:
`read.table` reads data into a data frame
`attach` makes names of column variables available
`ts` produces a time series object
`aggregate` creates an aggregated series
`ts.plot` produces a time plot for one or more series
`window` extracts a subset of a time series
`time` extracts the time from a time series object
`ts.intersect` creates the intersection of one or more time series
`cycle` returns the season for each value in a series trend, seasonal effect, and residual
`stl` decomposes a series using loess smoothing
`summary` summarises an R object

###Recursive functions: Factorial 

```{r}
#Calculate factorial
Fact <- function(n) if (n == 1) 1 else n * Fact(n - 1)
Fact(5)
```

```{r AirPassengers}
#data(AirPassengers)
AP <- AirPassengers
AP
```

```{r}
class(AP)
start(AP); end(AP); frequency(AP)
```
```{r}
summary(AP)
```

```{r}
plot(AP, ylab = "Passengers (1000's)")
```

## Seasonal Effect
To get a clearer view of the trend, the **seasonal effect can be removed** by
aggregating the data to the annual level, which can be achieved in R using the `aggregate()` function. 
A `summary()` of the values for each season can be viewed
using a `boxplot()`, with the `cycle()` function being used to **extract the seasons for each item of data**.
```{r}
layout(1:1)
plot(aggregate(AP))
boxplot(AP ~ cycle(AP))
```

```{r}
Maine.month <- read.delim("../../Data/Maine.dat", header=TRUE)

#attach() allows to access the columns in the Maine.month dataset directly
attach(Maine.month)
class(Maine.month)
```

```{r}
#now the unemploy field can be accessed directly
Maine.month.ts <- ts(unemploy, start = c(1996, 1), frequency = 12)
```
To select a smaller number of data we can specify an earlier end date using the parameter `end`. 

To analyse trends in the unemployment rate we can divide by 12 to give a mean annual rate.
```{r}
Maine.annual.ts <- aggregate(Maine.month.ts)/12
```
We now plot both time series. There is clear monthly variation. From
Figure 1.3(a) it seems that the February figure is typically about 20% more
than the annual average, whereas the August figure tends to be roughly 20%
less.

```{r}
layout(1:2)
plot(Maine.month.ts, ylab = "unemployed (%)")
plot(Maine.annual.ts, ylab = "unemployed (%)")
```

We can calculate the precise percentages in R, using **window**. This function will **extract that part of the time series between specified start and end points**
```{r}

Maine.Feb <- window(Maine.month.ts, start = c(1996,2), freq = TRUE)
Maine.Feb
```

```{r}
Feb.ratio <- mean(Maine.Feb) / mean(Maine.month.ts)
Feb.ratio
```

```{r}
layout(1:1)
plot(Maine.Feb)
```

```{r}
x <- c(1,2,3)
#Create a matrix col=3
y <- matrix(c(3,1,2,2,1,3), ncol = 3, byrow = TRUE)
y
```


```{r}
#Transpose the y matrix
t(y)
```
```{r}
#Multiplication: x[1,3]*y[3,2]=z[1,2]
x%*%t(y)
```
```{r}
t <- seq(from = 1, to = 100, by = 1) + 10 + rnorm(100, sd = 7)
time_series<- ts(t, start=2004, frequency = 4)
plot(time_series)
```
```{r}
#www <- "http://www.massey.ac.nz/~pscowper/ts/USunemp.dat"
US.month <- read.delim("../../Data/USunemp.dat", header=TRUE)

attach(US.month)
US.month.ts <- ts(US.month, start=c(1996,1), end=c(2006,10), frequency = 12)
plot(US.month.ts, ylab = "unemployed (%)")

```

The monthly supply of electricity (millions of kWh), beer (Ml),
and chocolate-based production (tonnes) in Australia over the period January
1958 to December 1990 are available from the Australian Bureau of Statistics
(ABS).

```{r}
#www <- "http://www.massey.ac.nz/~pscowper/ts/cbe.dat"
CBE <- read.delim("../../Data/cbe.dat", header = T)
CBE[1:4, ]

class(CBE)
```
create time series objects for the electricity, beer, and chocolate data.
If you omit end, R uses the full length of the vector, and if you omit the month
in start, R assumes 1. You can use plot with cbind to plot several series on
one figure (Fig. 1.5).

```{r}
Elec.ts <- ts(CBE[, 3], start = 1958, frequency =  12)
Beer.ts <- ts(CBE[, 2], start = 1958, frequency =  12)
Choc.ts <- ts(CBE[, 1], start = 1958, frequency =  12)
plot(cbind(Elec.ts, Beer.ts, Choc.ts), main="Chocolate, Beer, and Electricity Production: 1958-1990")
```
`ts.intersect` finction to obtain the intersection of two series that overlap in time.

```{r}
AP.elec <- ts.intersect(AP, Elec.ts)
start(AP.elec)
#[1] 1958 1
```

```{r}
end(AP.elec)
#[1] 1960 12
```

```{r}
AP.elec[1:3, ]

```

```{r}
AP <- AP.elec[,1]; Elec <- AP.elec[,2]
layout(1:2)
plot(AP, main = "", ylab = "Air passengers / 1000's")
plot(Elec, main = "", ylab = "Electricity production / MkWh")
```

```{r}

#as.vector() is needed to convert ts to ordinary vector required for a plot()

plot(as.vector(AP), as.vector(Elec),xlab = "Air passengers / 1000's",ylab = "Electricity production / MWh")
abline(reg = lm(Elec ~ AP))
```
##High correlation

```{r}
cor(AP, Elec)

```
In the plot function above, as.vector is needed to convert the ts objects to
ordinary vectors suitable for a scatter plot.

```{r}
#www <- "http://www.massey.ac.nz/~pscowper/ts/pounds_nz.dat"
Z <- read.delim("../../Data/pounds_nz.dat", header = T)
Z[1:4, ]
#[1] 2.92 2.94 3.17 3.25

```

```{r}
#quarterly: frequency = 4
Z.ts <- ts(Z, start = 1991, frequency = 4)

plot(Z.ts, xlab = "time / years",
ylab = "Quarterly exchange rate in $NZ / pound")
```
Two local trends are emphasised when the series is partitioned into two
subseries based on the periods 1992-1996 and 1996-1998. The window function
can be used to extract the subseries:


```{r}
Z.92.96 <- window(Z.ts, start = c(1992, 1), end = c(1996, 1))
Z.96.98 <- window(Z.ts, start = c(1996, 1), end = c(1998, 1))
layout (1:2)
plot(Z.92.96, ylab = "Exchange rate in $NZ/pound", xlab = "Time (years)" )
plot(Z.96.98, ylab = "Exchange rate in $NZ/pound", xlab = "Time (years)" )

```

In climate change studies (e.g., see Jones and Moberg, 2003; Rayner et al.
2003), the following global temperature series, expressed as anomalies from
the monthly means over the period 1961-1990, plays a central role
```{r}
#www <- "global.dat"
#Global <- scan(www)
#Global.ts <- ts(Global, start = c(1856, 1), end = c(2005, 12),frequency =  12)
Choc.annual <- aggregate(Choc.ts, FUN = mean)
plot(Choc.ts)
plot(Choc.annual)

```
using the `time()` function and the associated observed temperature series extracted using `window()`. The data are plotted and a line superimposed using a regression of temperature on the new time index 

```{r}
New.series <- window(Choc.ts, start=c(1970, 1), end=c(1985, 12))
New.time <- time(New.series)
plot(New.series); abline(reg=lm(New.series ~ New.time))
```
