---
title: "Matrix Operation"
author: "Helen Yezerets"
date: "March 2019""
output:
  html_document:
---



## Create Matrix using cbind, rbind, nrow, ncol, byrow=TRUE/FALSE
##dim() checks dimentions of the matrix
Name columns and rows

```{r }
# 3 x 2
x <- as.matrix(cbind(col1=c(1, 2, 3), col2=c(.5,.6,.7)))
x
dim(x)

# 2 x 3
y <- as.matrix(rbind(row1=c(1, 2, 3), row2=c(.5,.6,.7)))
y
dim(y)

#3 x 2
zR <- matrix(c(1, 2, 3, .5,.6,.7), nrow=3, byrow=TRUE)
zR
dim(zR)

#2 x 3
zC <- matrix(c(1, 2, 3, .5,.6,.7), ncol=3, byrow=FALSE)
zC
dim(zC)

```

##Create matrix using data.frame

```{r }
zD <- as.matrix(data.frame(ColName1 = c(1, 2, 3), 
                          ColName2 = c(.5,.6,.7)))
zD
```
## Matrix Algebra
### Addition of matrices

```{r }
x + zR
```
### Multiplication by scalar c
```{r }
c = 2
x * 2

```

### Matrix Multiplication uses `%*% ` and requires same inner parameters. It produces a matrix with outer parameter dimention: y(2,3) %*% x(3,2)= yx(2,2)

```{r}
y %*% x

```

### Matrix Transpose with `t()`

```{r}

zT <- t(x)
zT

```

## Matrix Metadata

```{r}
#Number of cols and rows
x
nrow(x)
ncol(x)
```

```{r}
#Sum of cols or rows
x
colSums(x)
rowSums(x)
```

```{r}
x
#Sum and mean of all elements
sum(x); mean(x)
```

```{r}
#Mean value by column or row
x
colMeans(x)
rowMeans(x)


```
#Quiz examples:
Create a 3x2 matrix m with the following values
0.85  1.5
0.75  1.0
0.65  0.5
Multiply the matrix m by the scalar c = 2
Find the mean of updated matrix. 

```{r}
m <- as.matrix(cbind(col1=c(.85,.75,.65), col2=c(1.5,1,.5)))
c=2
mean(m*c)
```

Transpose the following 3x2 matrix m
0.85  1.5
0.75  1.0
0.65  0.5
Find the means of columns with colMeans()
```{r}
t(m)
colMeans(t(m))
```