---
title: "ARIMA_modeling"
author: "Helen Yezerets"
date: "March 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Source: 
Chatterjee, S. (2018, February 05). Time Series Analysis Using ARIMA Model In R. Retrieved March 25, 2019, from https://datascienceplus.com/time-series-analysis-using-arima-model-in-r/

##What is ARIMA (AutoRegressive Integrated Moving Average) modeling
AR: lags of the differenced series
I : number of difference used to make the time series stationary
MA: lags of errors

##Assumptions of ARIMA model
1. Does not depend on time (stationary).Example: white noise series and series with cyclic behavior.
2. Univariate data (one variable). Auto-regression works with the past values.

```{r}
#Use astsa package
library(astsa)
data()                     # use this command to view all the loaded data 
```

```{r birth}
#Convert dataset to ts using monthly data
tsData = ts(birth, frequency = 12)
components.ts = decompose(tsData)
plot(components.ts)
```
