---
title: "FinalProject_fmri_ts_analysis"
author: "Helen Yezerets"
date: "April 24, 2019"
output: html_document

references: 
Chatterjee, S. (2018). Time Series Analysis Using ARIMA Model In R. Retrieved March 25, 2019, from https://datascienceplus.com/time-series-analysis-using-arima-model-in-r/.

Coghlan, A. (2010). Using R for Time Series Analysis. Retrieved April 24, 2019, from https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html#time-series-analysis

Correlation matrix : A quick start guide to analyze, format and visualize a correlation matrix using R software. (n.d.). Retrieved April 25, 2019, from http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software#compute-correlation-matrix-in-r

Cowpertwait, P. S. P., & Metcalfe, A. V. (2009). Introductory Time Series with R.(Ch.6.6 ARMA/ARIMA Models) New York: Springer. https://doi.org/10.1007/978-0-387-88698-5. 

Shumway, R., & Stooffer, D. S. (2016). Time Series Analysis and Its Applications With R Examples (Fourth Edition).Ch.3. Springer.from http://www.stat.pitt.edu/stoffer/tsa4/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Data Description - Analysis of fMRI (astsa data package with complete data set)

##Description 

fMRI data consists of location and stimulus timing related to percieved pain. Stimulus was applied for 32 seconds with 32 seconds waiting period. The sampling rate was one measurement of BOLD (blood oxigenation-level dependent) intensity every 2 seconds for 256 seconds (n = 128). The number of subjects under each condition varies.

##Details

The signal was measured in 9 brain locations (L ocation):
*[1] Cortex 1: Primary Somatosensory, Contralateral, 
*[2] Cortex 2: Primary Somatosensory, Ipsilateral, 
*[3] Cortex 3: Secondary Somatosensory, Contralateral, 
*[4] Cortex 4: Secondary Somatosensory, Ipsilateral, 
*[5] Caudate, 
*[6] Thalamus 1: Contralateral, 
*[7] Thalamus 2: Ipsilateral, 
*[8] Cerebellum 1: Contralateral and 
*[9] Cerebellum 2: Ipsilateral.

The stimuli (T reatment) with the number of respective sujects:
*[1] Awake-Brush (5 subjects), 
*[2] Awake-Heat (4 subjects), 
*[3] Awake-Shock (5 subjects), 
*[4] Low-Brush (3 subjects), 
*[5] Low-Heat (5 subjects), and 
*[6] Low-Shock (4 subjects). 

Example: fmri$L1T6 (as well as fmri[[6]]) (Location 1, Treatment 6) will show the data for Cortex 1: Primary Somatosensory, Contralateral location with Low-Shock (4 subjects) treatment
```{r}
#Need astsa package
#install.packages("astsa")
library(astsa) 
```

##variable description and its statistical summary: mean, sd, min, max

```{r fmri}
#Explore fmri data
str(fmri)
```

```{r}

#show first 6 lines of the L1T1 column
head(fmri$L1T1)

#show last 6 lines of the L1T1 column
tail(fmri$L1T1)
```

```{r}
#check if the data is time series
is.ts(fmri)
```

```{r}

#view the times
time(fmri)
```
```{r}
#Column summary for individual subjects
summary(fmri$L1T2)

```
## ts description: start, end, frequency

### time() function calculates a vector of time indices, with one element for each time index on which the series was observed.
```{r}
 time(fmri_AHeat1)
```

### deltat() function returns the fixed time interval between observations 

```{r}
deltat(fmri_AHeat1)
```
### frequency() function returns the number of observations per unit time. 
```{r}
 frequency(fmri_AHeat1)
```

#### cycle() function returns the position in the cycle of each observation.
```{r}

 cycle(fmri_AHeat1) 
 
```

```{r}

par(mfrow=c(3,2))
#Explore T2 on all of the cortex zones
ts.plot(as.ts(fmri$L1T2), col=3:7, main="Cortex 1: Primary Somatosensory, Contralateral", ylab="BOLD", xlab="Time (1 pt = 2sec)")
ts.plot(as.ts(fmri$L2T2), col=3:7, main="Cortex 2: Primary Somatosensory, Ipsilateral", ylab="BOLD", xlab="Time (1 pt = 2sec)")
ts.plot(as.ts(fmri$L3T2),col=3:7, main="Cortex 3: Secondary Somatosensory, Contralateral", ylab="BOLD", xlab="Time (1 pt = 2sec)")
ts.plot(as.ts(fmri$L4T2),col=3:7, main="Cortex 4: Secondary Somatosensory, Ipsilateral", ylab="BOLD", xlab="Time (1 pt = 2sec)")

ts.plot(as.ts(fmri$L8T2),col=3:7, main="Cerebellum 1: Contralateral", ylab="BOLD", xlab="Time (1 pt = 2sec)")
ts.plot(as.ts(fmri$L9T2),col=3:7, main="Cerebellum 2: Ipsilateral", ylab="BOLD", xlab="Time (1 pt = 2sec)")
```

Results: 
In response to Awake-Heat stimulation 
* Both Contralateral Cortex zone show potential strong correlation and cyclic increase in BOLD response in 16, 45, 77 and 110 sec that resembles a gradual addaptation to the stimulus. 
* On the other hand, the Primary Ipsilateral zone reacts with a steady increase, with a max at the 32 sec of the stimulus during the first application, however the second stimulus produces two picks: 32 sec and 64. 
* The secondary Ipsilateral zone follows the Contralateral zone pattern with higher variability of the BOLD response signal
* Cerebellum exibits response to the stimulus similar to the Cortex Secondary zone with slightly less pronounced cyclic variations
```{r}
#Explore the basal brain zone
par(mfrow=c(3,1))
ts.plot(as.ts(fmri$L6T2),col=3:7, main="Thalamus 1: Contralateral", ylab="BOLD", xlab="Time (1 pt = 2sec)")
ts.plot(as.ts(fmri$L7T2),col=3:7, main="Thalamus 2: Ipsilateral", ylab="BOLD", xlab="Time (1 pt = 2sec)")
ts.plot(as.ts(fmri$L5T2),col=3:7, main="Caudate", ylab="BOLD", xlab="Time (1 pt = 2sec)")

```
The basal structures appear to respond as a WN. Need further investigation

```{r}
#Check row means
rowMeans(fmri$L1T2)
```


```{r}
#combine brain zones per stimulus using row means

#[1] Awake-Brush (5 subjects)
fmri_ABrush<-cbind(rowMeans(fmri$L1T1), rowMeans(fmri$L2T1), rowMeans(fmri$L3T1), rowMeans(fmri$L4T1), rowMeans(fmri$L5T1),  rowMeans(fmri$L6T1), rowMeans(fmri$L7T1), rowMeans(fmri$L8T1), rowMeans(fmri$L9T1))

#[2] Awake-Heat (4 subjects)
fmri_AHeat<-cbind(rowMeans(fmri$L1T2), rowMeans(fmri$L2T2), rowMeans(fmri$L3T2), rowMeans(fmri$L4T2), rowMeans(fmri$L5T2),  rowMeans(fmri$L6T2), rowMeans(fmri$L7T2), rowMeans(fmri$L8T2), rowMeans(fmri$L9T2))
#add column names
colnames(fmri_AHeat) <-c("pSS(C)", "pSS(I)","sSS(C)", "sSS(I)", "Cau", "Tha(C)","Tha(I)", "Crb(C)", "Crb(I)")


#[3] Awake-Shock (5 subjects), 
fmri_AShock<-cbind(rowMeans(fmri$L1T3), rowMeans(fmri$L2T3), rowMeans(fmri$L3T3), rowMeans(fmri$L4T3), rowMeans(fmri$L5T3),  rowMeans(fmri$L6T3), rowMeans(fmri$L7T3), rowMeans(fmri$L8T3), rowMeans(fmri$L9T3))
#add column names for Sedated Heat
colnames(fmri_AShock) <-c("pSS(C)", "pSS(I)","sSS(C)", "sSS(I)", "Cau", "Tha(C)","Tha(I)", "Crb(C)", "Crb(I)")

#[4] Low-Brush (3 subjects),
fmri_SBrush<-cbind(rowMeans(fmri$L1T4), rowMeans(fmri$L2T4), rowMeans(fmri$L3T4), rowMeans(fmri$L4T4), rowMeans(fmri$L5T4),  rowMeans(fmri$L6T4), rowMeans(fmri$L7T4), rowMeans(fmri$L8T4), rowMeans(fmri$L9T4))

#[5] Low-Heat (5 subjects),  
fmri_SHeat<-cbind(rowMeans(fmri$L1T5), rowMeans(fmri$L2T5), rowMeans(fmri$L3T5), rowMeans(fmri$L4T5), rowMeans(fmri$L5T5),  rowMeans(fmri$L6T5), rowMeans(fmri$L7T5), rowMeans(fmri$L8T5), rowMeans(fmri$L9T5))
#add column names for Sedated Heat
colnames(fmri_SHeat) <-c("pSS(C)", "pSS(I)","sSS(C)", "sSS(I)", "Cau", "Tha(C)","Tha(I)", "Crb(C)", "Crb(I)")


#[6] Low-Shock (4 subjects). 
fmri_SShock<-cbind(rowMeans(fmri$L1T6), rowMeans(fmri$L2T6), rowMeans(fmri$L3T6), rowMeans(fmri$L4T6), rowMeans(fmri$L5T6),  rowMeans(fmri$L6T6), rowMeans(fmri$L7T6), rowMeans(fmri$L8T6), rowMeans(fmri$L9T6))
#add column names for Sedated Heat
colnames(fmri_SShock) <-c("pSS(C)", "pSS(I)","sSS(C)", "sSS(I)", "Cau", "Tha(C)","Tha(I)", "Crb(C)", "Crb(I)")

```
##Depiction of the  response to Heat by various brain zones
### Awake
```{r}

summary(fmri_AHeat)

plot.ts(fmri_AHeat,col=4,xlab="Time (1 pt = 2sec)", main="Awake Heat")

#Show subtitle on a side of the plot (1=bottom, 2=left, 3=top, 4=right)
mtext("Cortex & Caudate", side=3, adj=.05, line=-0.8) 
mtext("Thalamus & Cerebellum", side=3, adj=.95, line=-0.8)

```


##Compare primary and secondary SS(C) and SS(I)
```{r}
#fmri_AHeat[,2]
summary(fit<-lm(fmri_AHeat[,2]~fmri_AHeat[,1]))
plot(fmri_AHeat[,1],fmri_AHeat[,2])
abline(fit, col=2)
```
```{r}
#Check correlation
cor(fmri_AHeat[,1],fmri_AHeat[,2],method=c("pearson","kendall","spearman"))
```
##Let's create a correlation matrix for Awake-Heat data to compare locations 
```{r}
#Install Hmisc package:
#install.packages("Hmisc")
#Use rcorr() function
library("Hmisc")
```

Use corrplot() function to draw a correlogram. The function corrplot(), in the package of the same name, creates a graphical display of a correlation matrix, highlighting the most correlated variables in a data table.

```{r}
#Install corrplot
#install.packages("corrplot")
library(corrplot)
```

Correlation coefficients are colored according to the value. Correlation matrix can be also reordered according to the degree of association between variables.
The function corrplot() takes the correlation matrix as the first argument. The second argument (type="upper") is used to display only the upper triangular of the correlation matrix.

#### Let's combine correlogram with the significance test. We'll use the result generated  with rcorr() function [in Hmisc package]
Insignificant correlations are leaved blank
Order method of the correlation matrix ("original" for original order (default), "AOE" for the angular order of the eigenvectors, "FPC" for the first principal component order, "hclust" for the hierarchical clustering order, "alphabet" for alphabetical order)
```{r}
#For Awake Heat 
ah2 <- rcorr(as.matrix(fmri_AHeat))
ah2

corrplot(ah2$r,method="square", type="upper", 
          p.mat = ah2$P, sig.level = 0.01, insig = "blank",
         tl.col="black")
```


###Sedated Heat
```{r}
summary(fmri_SHeat)

plot.ts(fmri_SHeat,col=4,xlab="Time (1 pt = 2sec)", main="Sedated Heat")

#Show subtitles 
mtext("Cortex & Caudate", side=3, adj=.05, line=-0.8) 
mtext("Thalamus & Cerebellum", side=3, adj=.95, line=-0.8)

```
####Correlation Martix for the Sedated Heat stimulus

```{r}
#For Sedated Heat 
sh2 <- rcorr(as.matrix(fmri_SHeat))
#sh2

corrplot(sh2$r,method="square", type="upper", 
          p.mat = sh2$P, sig.level = 0.01, insig = "blank",
         tl.col="black")
```

## Depiction of the response to Shock
### Awake
```{r}

summary(fmri_AShock)

plot.ts(fmri_AShock,col=4,xlab="Time (1 pt = 2sec)", main="Awake Shock")

#Show subtitles 
mtext("Cortex & Caudate", side=3, adj=.05, line=-0.8) 
mtext("Thalamus & Cerebellum", side=3, adj=.95, line=-0.8)

```
Results:similar response to heat. More pronounced sinusoid shape of the response in pSSC and Crb(I) that closely resemble each other.



####Correlation Martix for the Awake Shock stimulus

```{r}
as2 <- rcorr(as.matrix(fmri_AShock))
#as2

corrplot(as2$r,method="square", type="upper", 
          p.mat = as2$P, sig.level = 0.01, insig = "blank",
         tl.col="black")
```
### Sedated
```{r}
summary(fmri_SShock)

plot.ts(fmri_SShock,col=4,xlab="Time (1 pt = 2sec)", main="Sedated Shock")

#Show subtitles 
mtext("Cortex & Caudate", side=3, adj=.05, line=-0.8) 
mtext("Thalamus & Cerebellum", side=3, adj=.95, line=-0.8)

```
Results: WN? while sedated

####Correlation Martix for the Sedated Shock stimulus

```{r}
ss2 <- rcorr(as.matrix(fmri_SShock))
#ss2

corrplot(ss2$r,method="square", type="upper", 
          p.mat = ss2$P, sig.level = 0.01, insig = "blank",
         tl.col="black")
```

```{r}
#ts.plot(as.ts(fmri_AShock[,1]), as.ts(fmri_SShock[,1]),col=3:4,xlab="Time (1 pt = 2sec)", main="Shock")
#legend('topright', legend=c("Awake","Sedated"), lty=1, col=3:4)
```

###Decompose Awake Heat data based on the combined locations with the highest significance for correlation 
To use decompose function the frequency has to equal at least 2. Decompose() function makes seasonal adjustments using moving average, depicts original ts, trend, seasonal, and random (remainder) part

```{r}
# pSS, sSS, Crb(C)
fmri_AHeat1<-ts(c(fmri_AHeat[,1],fmri_AHeat[,2],fmri_AHeat[,3],fmri_AHeat[,4],fmri_AHeat[,8]), start=1, end=64, frequency=2)
#is.ts(fmri_AHeat1)
#fmri_AHeat1
m1<-decompose(fmri_AHeat1)
plot(ts(m1$random[1:128]))
acf(m1$random[1:128], na.action=na.pass) #plots autocovariance or autocorrelation
```

```{r}
#Check for NA values
m1$random[1:128]
```

# Calculate st dev to check if the decomposition made any effect
```{r}
sd(fmri_AHeat1)
sd(fmri_AHeat1[2:126] - m1$trend[2:126])
sd(m1$random[2:126])

```
Results: The reduction in the standard deviation shows that the seasonal adjustment has been effective.



###Decompose Sedated Heat data based on the combined locations with the highest significance for correlation 

```{r}
# Caudate, Thalamus and Cerebellum
fmri_SHeat1<-ts(c(fmri_SHeat[,5],fmri_SHeat[,6],fmri_SHeat[,7],fmri_SHeat[,8],fmri_SHeat[,9]), start=1, end=64, frequency=2)

m2<-decompose(fmri_SHeat1)
plot(ts(m2$random[1:128]))
acf(m2$random[1:128], na.action=na.pass) #plots autocovariance or autocorrelation
```

# Calculate st dev to check if the decomposition made any effect
```{r}
sd(fmri_SHeat1)
sd(fmri_SHeat1[2:126] - m2$trend[2:126])
sd(m2$random[2:126])

```
Results: The reduction in the standard deviation shows that the seasonal adjustment has been effective.

###Decompose Awake Shock data based on the combined locations with the highest significance for correlation 

```{r}
# sSS + Crb(I)
fmri_AShock1<-ts(c(fmri_AShock[,3],fmri_AShock[,4],fmri_AShock[,9]), start=1, end=64, frequency=2)

m3<-decompose(fmri_AShock1)
plot(ts(m3$random[1:128]))
acf(m3$random[1:128], na.action=na.pass) #plots autocovariance or autocorrelation
```

# Calculate st dev to check if the decomposition made any effect
```{r}
sd(fmri_AShock1)
sd(fmri_AShock1[2:126] - m3$trend[2:126])
sd(m3$random[2:126])

```
Results: The reduction in the standard deviation shows that the seasonal adjustment has been effective.

###Decompose Sedated Shock data based on the combined locations 

```{r}
#pSS
fmri_SShock1<-ts(c(fmri_SShock[,1],fmri_AShock[,2]), start=1, end=64, frequency=2)

m4<-decompose(fmri_SShock1)
plot(ts(m4$random[1:128]))
acf(m4$random[1:128], na.action=na.pass) #plots autocovariance or autocorrelation
```

# Calculate st dev to check if the decomposition made any effect
```{r}
sd(fmri_SShock1)
sd(fmri_SShock1[2:126] - m4$trend[2:126])
sd(m4$random[2:126])

```
Results: The slight reduction in the standard deviation shows that the seasonal adjustment has been only mildly effective.

#Let's check if it is White Noise.

## WN assumptions:
### 95% of values ~= 0  and lay inside the blue lines
### They should be uncorrelated
### They should have mean zero
Useful properties (for computing prediction intervals)
### They should have constant variance
### They should be normally distributed

#Use Ljung-Box test to check for WN
The Ljung-Box test considers the first h autocorrelation values together.

```{r}
Box.test(fmri_SShock1, lag = 16, fitdf = 0, type = "Lj")

```
p-value = 0.01828 indicates the data are probably White Noise.

#Summary: 
We can test for white noise by looking at an ACF plot or by doing a Ljung-Box test

#Forecasts and potential futures
The 80% forecast intervals should contain 80% of the future observations
The 95% forecast intervals should contain 95% of the future observations




##Reduce seasonal effect by aggregating data using aggregate() and cycle() . 
```{r message=FALSE, echo=FALSE}
layout(1:2)
plot(aggregate(fmri_AHeat1), ylab = "Aggregated")
boxplot(fmri_AHeat1 ~ cycle(fmri_AHeat1))
```
##Reduce seasonal effect by aggregating data using aggregate() and cycle() . 
```{r message=FALSE, echo=FALSE}
layout(1:2)
plot(aggregate(fmri_AHeat2), ylab = "Aggregated")
boxplot(fmri_AHeat2 ~ cycle(fmri_AHeat2))
```

##Reduce seasonal effect by aggregating data using aggregate() and cycle() . 
```{r message=FALSE, echo=FALSE}
layout(1:2)
plot(aggregate(fmri_AHeat3), ylab = "Aggregated")
boxplot(fmri_AHeat3 ~ cycle(fmri_AHeat3))
```