---
title: "datacamp_xts3_merging_ts"
author: "Helen Yezerets"
date: "February 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Combining xts by column with merge
**xts** makes it easy to join data by column and row using a few different functions. All results will be correctly ordered in time, regardless of original frequencies or date class. One of the most important functions to accomplish this is **merge()**. It takes one or more series and joins them by column. It's also possible to combine a series with a vector of dates. This is especially useful for normalizing observations to a fixed calendar.

merge() takes three key arguments:
* an arbitrary number of objects to combine. 
* **join** (inner, left or right,  default join = "outer"). This is similar to a relational database join, only here, the index is what we join on. 
* **fill** - controls missingness introduced as a result of the merge, ex. **fill=na.locf** (last observation carried forward)

# Basic argument use
merge(a, b, join = "right", fill = 9999)
```{r}
# Perform an inner join of a and b
merge(a, b, join = "inner")

# Perform a left-join of a and b, fill missing values with 0
merge(a, b, join = "left", fill = 0)

```

##Combining xts by row with rbind
How to add new rows to your data.

xts provides its own S3 method to the base rbind() generic function. The xts rbind function is much simpler than merge(). The only argument that matters is ..., which takes an arbitrary number of objects to bind. What is different is that rbind requires a time series, since we need to have timestamps for R to know where to insert new data.

```{r}
# Row bind temps_june30 to temps, assign this to temps2
temps2 <- rbind(temps_june30, temps)

# Row bind temps_july17 and temps_july18 to temps2, call this temps3
temps3 <- rbind(temps_july17, temps_july18, temps2)

```
Because xts objects are ordered by their time index, the order of arguments in xts's rbind() command is unimportant.

##What types of data can be combined using merge?
Although xts is very flexible when it comes to binding new columns to existing objects, there are still some exceptions.

Answer:
xts objects of identical type (e.g. integer + integer).
POSIXct dates vector.
Atomic vectors of the same type (e.g. numeric).

No merge for data.frames with various column types.

##Fill missing values using last or previous observation
The xts package leverages the power of zoo for help with missingness. Example: **na.locf()** - last observation carried forward approach. It both preserves the last known value and prevents any look-ahead bias from entering into the data.

You can also apply next observation carried backward by setting **fromLast = TRUE**.

# Last obs. carried forward
na.locf(x)                

# Next obs. carried backward
na.locf(x, fromLast = TRUE) 
 
# Fill missing values in temps using the last observation
temps_last<-na.locf(temps)

# Fill missing values in temps using the next observation
temps_next<-na.locf(temps, fromLast = TRUE) 

##NA interpolation using na.approx()
* It may be that a series is missing an observation due to a higher frequency sampling than the generating process. 
* You might also encounter an observation that is in error, yet expected to be somewhere between the values of its neighboring observations.

These are scenarios where interpolation is useful. zoo provides a powerful tool to do this. Based on simple linear interpolation between points, implemented with **na.approx()** the data points are approximated using the distance between the index values. In other words, the estimated value is linear in time.

```{r}
# Interpolate NAs using linear approximation
na.approx(AirPass)
```

##Combine a leading and lagging time series
Another common modification for time series is the ability to lag a series. Also known as a **backshift operation**, it's typically shown in literature using Lk notation, indicating a transformation in time LkX=Xt???k. This lets you see observations like yesterday's value in the context of today.

Both zoo and xts implement this behavior, and in fact extend it from the ts original in R. There are two major differences between xts and zoo implementations that you need to be aware of. One is the direction of the lag for a given k. The second is how missingness is handled afterwards.

For historical reasons in R, zoo uses a convention for the sign of k in which negative values indicate lags and positive values indicate leads. That is, in zoo lag(x, k = 1) will shift future values one step back in time. This is inconsistent with the vast majority of the time series literature, but is consistent with behavior in base R. xts implements the exact opposite, namely for a positive k, the series will shift the last value in time one period forward; this is consistent with intuition, but quite different than zoo.

In this exercise, you will construct a single xts object with three columns. The first column is data one day ahead, the second column is the original data, and the third column is the one day behind - all using xts. A simple xts object, x, has been loaded into your workspace.

# Your final object
cbind(lead_x, x, lag_x)

# Create a leading object called lead_x
lead_x <- lag(x, k = -1)

# Create a lagging object called lag_x
lag_x<-lag(x, k=1)

# Merge your three series together and assign to z
z<-merge(lead_x,x, lag_x)

##Calculate a difference of a series using diff()
Another common operation on time series, typically on those that are non-stationary, is to take a difference of the series. The number of differences to take of a series is an application of recursively calling the difference function n times.

A simple way to view a single (or "first order") difference is to see it as x(t) - x(t-k) where k is the number of lags to go back. Higher order differences are simply the reapplication of a difference to each prior result.

In R, the difference operator for xts is made available using the diff() command. This function takes two arguments of note. The first is the lag, which is the number of periods, and the second is differences, which is the order of the difference (e.g. how many times diff() is called).

# These are the same
diff(x, differences = 2)
diff(diff(x))
# Calculate the first difference of AirPass and assign to diff_by_hand
diff_by_hand <- AirPass - lag(AirPass)

# Use merge to compare the first parts of diff_by_hand and diff(AirPass)
merge(head(diff_by_hand), head(diff(AirPass)))

# Calculate the first order (how many times it is called) 12 month (lag=12) difference of AirPass
diff(AirPass, lag = 12, defferences = 1)

##What is the key difference in lag between xts and zoo
As you've seen, generating lags and leads are an important tool in your arsenal for handling time series data in R. However, because xts employs slightly different procedures for generating lags and leads compared with zoo of base-R, you have to be very precise in your calls to lag().

The k argument in zoo uses positive values for shifting past observations forward. - FALSE

The k argument in xts uses positive values for shifting past observations forward. -TRUE