---
title: "datacamp_xts4_functions"
author: "Helen Yezerets"
date: "February 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Find intervals by time in xts
**endpoints()**: input(a vector of times) and output(the locations of the last observations in each interval).

For example, the code below locates the last observation of each year for the AirPass data set.

endpoints(AirPass, on = "years")
[1] 0 12 24 36 48 60 72 84 96 108 120 132 144

The argument **on** supports a variety of periods, including **"years", "quarters", "months", as well as intraday intervals such as "hours", and "minutes"**. What is returned is a vector starting with 0 and ending with the extent (last row) of your data.

In addition to each period, you can find the Kth period by utilizing the k argument. For example, setting the arguments of your **endpoints() call to on = "weeks", k = 2**, would generate the final day of every other week in your data. 
###Note
*Always* starts on 0 and ends on the last observation
in other words, **last value returned will always be the length of the input data**, even if it doesn't correspond to a skipped interval.

```{r}
library(xts)
data(AirPassengers)

#Convert to xts object
AirPass<-as.xts(AirPassengers)

# Locate locates the last observation of each year 
endpoints(AirPass, on = "years")

# Locate locates the last observation of the third (k=3) year 
endpoints(AirPass, on = "years", k = 3)
```

##Apply a function by time period(s)
We can subset the object to identify the last known value. For most series, apply a function to the values between endpoints. In essence, use the base function **apply()**, but used on a window of time with **period.apply()** command, which takes a time series (x), an index of endpoints, and a function:
**period.apply(x, INDEX, FUN, ...)**

```{r}
# Calculate the yearly endpoints
ep <- endpoints(AirPass, on = "years", k=2)

# Now calculate the yearly mean and display the results
period.apply(AirPass, INDEX = ep, FUN = mean)

```
##Using lapply() and split() to apply functions on intervals
To physically split your data into disjoint chunks by time and perform some calculation on these periods use of the xts **split()** command to chunk your data by time. The split() function creates a list containing an element for each split. The **f** argument in split() is a character string **describing the period to split by (i.e. "months", "years", etc.)**.

The function **lapply()** is used for the most efficient calculations. In cases where you don't want to return a time series, this proves to be very intuitive and effective.

```{r}
# Split temps by year
air_yearly <- split(AirPass, f = "years", k=2)
air_yearly

# Create a list of yearly means, temps_avg, and print this list
air_avg <- lapply(air_yearly, FUN = mean)
air_avg
```
**period.apply()** is similar to using a combination of **split() and lapply()**.

##Selection by endpoints vs. split-lapply-rbind
Find the last observation in each year

**xts** represents **irregular time series**, so it is perfectly valid to have holes in the data where one might expect an observation.
```{r}
# Use the proper combination of split, lapply and rbind
x_1 <- do.call(rbind, lapply(split(AirPass, "years"), function(w) last(w, n = 1)))
#x_1
```

```{r}

# Create last_month_of_years using endpoints()
last_month_of_years <- endpoints(AirPass, "years")
last_month_of_years
```

```{r}
# Subset AirPass using last_month_of_years 
x_2 <- AirPass[last_month_of_years]
x_2
```

##Convert univariate series to Open-High-Low-Close (or OHLC) data
Aggregating time series can be a frustrating task. For example, in financial series it is common to find Open-High-Low-Close data (or OHLC) calculated over some repeating and regular interval.

Also known as range bars, aggregating a series based on some regular window can make analysis easier amongst series that have varying frequencies. A weekly economic series and a daily stock series can be compared more easily if the daily is converted to weekly.

To convert from a univariate series into OHLC series, and then convert your final OHLC series back into a univariate series using the xts function **to.period()** that takes a time-series, x, and a string for the period (i.e. months, days, etc.), in addition to a number of other optional arguments.

to.period(x,
          period = "months", 
          k = 1, 
          indexAt, 
          name=NULL,
          OHLC = TRUE,
          ...)

```{r}
# Convert usd_eur to weekly and assign to usd_eur_weekly
usd_eur_weekly <- to.period(usd_eur, period = "weeks")

# Convert usd_eur to monthly and assign to usd_eur_monthly
usd_eur_monthly <- to.period(usd_eur, period = "months")

# Convert usd_eur to yearly univariate and assign to usd_eur_yearly
usd_eur_yearly <- to.period(usd_eur, period = "years", OHLC = FALSE)

```

##Convert a series to a lower frequency
Besides converting univariate time series to OHLC series, **to.period() also lets you convert OHLC to lower regularized frequency** - something like subsampling your data.

Depending on the chosen frequency, the index class of your data may be coerced to something more appropriate to the new data. For example, when using the shortcut function **to.quarterly()**, xts will convert your index to the yearqtr class to make periods more obvious.

We can override this behavior by using the **indexAt** argument. Specifically, using **firstof** would give you the time from the beginning of the period. In addition, you can change the base name of each column by supplying a string to the argument name.

For this exercise we'll introduce a new dataset, the edhec hedge fund index data from the PerformanceAnalytics package.

In this exercise you will use the Equity Market Neutral time series from the edhec data, which we've assigned to eq_mkt.

```{r}
# Convert eq_mkt to quarterly OHLC
mkt_quarterly <- to.period(eq_mkt, period = "quarters")

# Convert the original eq_mkt again, this time using to.quarterly() directly. Change the base name of each OHLC column to edhec_equity and change the index to "firstof". Call this mkt_quarterly2.
mkt_quarterly2 <- to.quarterly(eq_mkt, name = "edhec_equity", indexAt = "firstof")
```

##Calculate basic rolling value of series by month
One common aggregation you may want to apply involves doing a calculation within the context of a period, but returning the interim results for each observation of the period.

For example, you may want to calculate a running month-to-date cumulative sum of a series. This would be relevant when looking at monthly performance of a mutual fund you are interested in investing in.

For this exercise, you'll calculate the cumulative annual return using the edhec fund data from the last exercise. To do this, you'll follow the split()-lapply()-rbind() pattern demonstrated below:

x_split <- **split(x, f = "months")**
x_list <- **lapply(x_split, cummax)**
x_list_rbind <- **do.call(rbind, x_list)**

Note the last call uses R's somewhat strange do.call(rbind, ...) syntax, which allows you **to pass a list to rbind** instead of passing each object one at a time. This is a handy shortcut.

```{r}
# Split edhec into years
edhec_years <- split(edhec , f = "years")

# Use lapply to calculate the cumsum for each year in edhec_years
edhec_ytd <- lapply(edhec_years, FUN = cumsum)

# Use do.call to rbind the results
edhec_xts <- do.call(rbind, edhec_ytd)

```
##Calculate the rolling standard deviation of a time series
Another common requirement when working with time series data is to apply a function on a rolling window of data. xts provides this facility through the intuitively named zoo function **rollapply()**.

This function takes a time series object x, a window size width, and a function FUN to apply to each rolling period. The **width argument** can be tricky; a number supplied to the width argument specifies the number of observations in a window. For instance, to take the **rolling 10-day max **of a series, you would type the following:

**rollapply(x, width = 10, FUN = max, na.rm = TRUE)**
Note that the above would only take the 10-day max of a series with daily observations. If the series had monthly observations, it would take the 10-month max. Also note that you can pass additional arguments (i.e. na.rm to the max function) just like you would with apply().
```{r}
# Use rollapply to calculate the rolling 3 period sd of eq_mkt
eq_sd <- rollapply(eq_mkt, 3, FUN = sd)

```


